---
title: "Spatial Analysis"
author: "José Augusto Mendonça"
date: "12/09/2021"
output: html_document
---

# 0. Set up R

```{r setup, include=FALSE}

# Clear workspace

rm(list = ls()) 

# Load packages for spatial analysis

library(sp)
library(rgdal)
library(rgeos)
library(ggmap)                 
library(leaflet)
library(spatstat)
library(gstat)
library(maptools)
library(ggplot2)
library(sf)
library(dplyr)

# Numbers do not display as scientific notation 

options(scipen = 999)

# Read GEOJSON file

library("geojsonio")

```

Geojson files are common way to store geospatial data.

# 1. Read data

```{r read_data}

# Load local R Dataframe

amsterdam <- read.csv('dfgeolocation.csv')

# Check the structure of the Dataframe

class(amsterdam)

# Check the structure of the Dataframe variables 

str(amsterdam)

# Check for missing values (NA´s)

summary(amsterdam)

# Spatial data

spdf <- geojson_read('ciclomapa.geojson')

```

# 2. Mutate data

```{r mutate_data, echo=FALSE}

# Remove missing values from Dataframe

# amsterdam <- filter(amsterdam,lat != "NA")

amsterdam <- amsterdam[!is.na(amsterdam$lat), ] 

# Mutate time variable as date format

amsterdam$registrationdate <- as.Date(amsterdam$registrationdate)
 
```

# 3. Spatial Points Dataframe

We need to convert *amsterdam* local Dataframe into *SpatialPointsDataFrame* using function from *sp* package

proj4string is the Coordinate Reference System.

By running the CRS function, we discarded unknown data based on WGS84 reference

```{r spatial_points}

# Set the coordinates object - Facilitate R to recognize that these two variables are coordinates

# Before R used to recognize them as numbers

coords <- SpatialPoints(amsterdam[, c("lon", "lat")])

# Create spatial Dataframe  

amsterdam_spatial_df <- SpatialPointsDataFrame(coords, amsterdam) 

# Define our type of coordinates reference system

proj4string(amsterdam_spatial_df) <- CRS("+proj=longlat +ellps=WGS84") 
 
```

# 4. Have a quick look on what we are doing

The foundation class is the *Spatial* class

```{r quicklook}

# See the possible classes for Class Spatial

getClass("Spatial")

# Check structure of new Dataframe (We converted from local to Spatial Dataframe)

class(amsterdam_spatial_df) 

# Plot the points in a plane

plot(amsterdam_spatial_df)
 
```

# 5. Save Spatial data

```{r savespatialdata}

# Function saveRDS - Maintain the Spatial integrity

saveRDS(amsterdam_spatial_df, "amsterdam_spatial_df.rds")

```

# 6. Load polygon shapefiles (geographical layer)

```{r load_polygon}

# Read shape files (First folder, then file)

# Country level

nl_shp <- readOGR(dsn = "C:/Users/jose_/OneDrive/Área de Trabalho/Spatial Test/NLD_adm", "NLD_adm0")

# Region level

nl1_shp <- readOGR(dsn = "C:/Users/jose_/OneDrive/Área de Trabalho/Spatial Test/NLD_adm", "NLD_adm1")

# Municipality level

nl2_shp <- readOGR(dsn = "C:/Users/jose_/OneDrive/Área de Trabalho/Spatial Test/NLD_adm", "NLD_adm2")

# Check the boundering plots

plot(nl_shp)
plot(nl1_shp)
plot(nl2_shp)
 
```

# 7. Spatial Descriptive Statistics

```{r descriptive}

# Check the bounding area of a spatial object

bbox(amsterdam_spatial_df)

# Explore the projection (coordination) system of any spatial object

proj4string(amsterdam_spatial_df)

# Have a look into the first observations

head(amsterdam_spatial_df@coords, 5)
 
```

# 8. Data visualization

Let's add housing data onto the map. We can use the argument, 'add=TRUE' to add another layer onto the plot. 

```{r visualization}

# Plot and add layers

plot(nl2_shp, col = "grey", axes = TRUE)
plot(amsterdam_spatial_df, pch = 21, bg = "red", cex = .5, add=TRUE)
title("Housings solds in the NL")
legend("topleft", title = "Legend", legend = "Sold houses locations", pch = 21, pt.bg = "red", bty = "n")
 
```

# 9. Data visualization using ggplot2

ggplot2 works with data frames and not objects of class **Spatial**. So we have to convert them using 'fortify()'

```{r ggplot2}

# Convert spatial objects into dataframes 

amsterdam_df <- data.frame(amsterdam_spatial_df)
nl2_shp_df <- fortify(nl2_shp)

# Set map object

p <- ggplot() +
  geom_polygon(aes(x=long, y = lat, group=group), data = nl2_shp_df, colour = "black", fill = NA) +
  geom_point(data = amsterdam_df, aes(x = lon, y = lat), color = ("red"), size=1) +
  coord_equal() + labs(title = "dale") + xlab("Longitude") + ylab("Latitude")

# Print map object

p

```

# 10. Other type of ggplot Plot

```{r gglplot4}

# Define a combined object by a common variable (In this case, latitude)

map_and_data <- inner_join(amsterdam_df, nl2_shp_df)

# Plot combined object

ggplot(map_and_data) +
  geom_sf(aes(fill= obj_hid_TRANSACTIEPRIJS)) +
  scale_fill_gradient(low = "#56B1F7", high = "#132B43")


```

# 11. Data visualization using ggmap

```{r ggmap}

# Allow API key

register_google(key = "")

# Create Static background layer for the city of Amsterdam 

map_ams <- get_map(location = "amsterdam", source = "osm", zoom = 14)

# Use ggmap command to make the plot

amsterdamMap <- ggmap(map_ams, extent = "device", legend = "topleft")
amsterdamMap

# Plot the shapefiles

amsterdamMap2 <- amsterdamMap +
  geom_polygon(aex(x = long, y = lat, group = group), data = nl2_shp_df, colour = "black", fill = NA)

amsterdamMap2
 
```

# End
